\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-07-31 tex/ms.tex
added recursion
added figs

%
\section{The Spot Expansion}
\label{sec:spot}
%
\subsection{Preliminaries}
%
We adopt the following expression for the spherical harmonic coefficient
of degree $l$ and order $m$ in the expansion of a spot at
$\theta = \varphi = 0$:
%
\begin{align}
    \label{eq:ylm0}
    y_{lm}(r, \delta) =
    \begin{cases}
        1 - \dfrac{\delta r'}{2 (1 + r')}
         & l = m = 0    \\[2em]
        -\dfrac{\delta r' \left( 2 + r' \right)}
        {2 \sqrt{2l + 1} (1 + r')^{l + 1}}
         & l > 0, m = 0 \\[2em]
        0
         & m \ne 0
    \end{cases}
\end{align}
%
where
%
\begin{align}
    \label{eq:rprime}
    r' \equiv c_0 + c_1 r
\end{align}
%
and $c_0, c_1 > 0$ are scaling constants (defined in more detail below).
In the expressions above,
$\delta \in (-\infty, 1]$ is the spot contrast (the fractional
decrease in the brightness
at the center of the spot) and $r \in [0, 1]$ is the normalized spot radius.
The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
five important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity is azimuthally symmetric about the spot center
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $1 - \delta$
    \item The surface intensity at the antipode of the spot center is unity
    \item The size of the spot increases monotonically with $r$
\end{enumerate}
%
These properties may be demonstrated by considering the expression for the
surface intensity at a given point $(\theta, \varphi)$:
%
\begin{align}
    \begin{split}
        I(\theta, \varphi)
        & =
        \sum\limits_{l=0}^\infty \sum\limits_{m=-l}^l
        y_{lm} Y_{lm}(\theta, \varphi) \\
        & = \sum\limits_{l=0}^\infty
        y_{l0} \sqrt{2l + 1} P_l(\cos\theta)
    \end{split}
\end{align}
%
where $P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Note that the expression above is independent of the azimuth $\varphi$,
a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
expansion is therefore azimuthally symmetric about the spot center, as
stated in \textbf{1}.
%
Combining the above expression with Equation~(\ref{eq:ylm0}) and
rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    I(\theta) =
    1 + \dfrac{\delta r'}{2}
    -
    \dfrac{\delta r' \left( 2 + r' \right)}{2 (1 + r')}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + r'}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
    %
    \\
    \intertext{where}
    %
    \begin{split}
        A & = 1 + \dfrac{\delta r'}{2}                      \\
        B & = \delta r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
        C & = \dfrac{1 + (1 + r')^2}{2 + 2 r'}
    \end{split}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{2}. The value at the minimum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
        & = 1 - \delta
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{3}, and the value at the maximum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
        & = 1
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{4}.
Finally, to show \textbf{5}, let us compute the half width at half minimum
$\Delta\theta$ of the intensity profile:
%
\begin{align}
    1 - \left(A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}}\right) =
    \dfrac{1}{2}\delta
\end{align}
%
Solving for $\Delta\theta$ yields
%
\begin{align}
    \Delta\theta =
    \cos^{-1} \left[ \dfrac{2 + 3 r' (2 + r')}{2 (1 + r')^3} \right]
    \quad.
\end{align}
%
Differentiation with respect to $r'$ yields
%
\begin{align}
    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r'} =
    \frac{3 r' \left(2 + r'\right)}{\left(1 + r'\right)^4
        \sqrt{\frac{r'^2 \left(2 + r'\right)^2 \left(1 + 2 r'\right)
                \left(3 + 2 r'\right)}{\left(1 + r'\right)^6}}}
    \quad,
\end{align}
%
which is positive definite for all $r' > 0$.

@@

    \dfrac{\alpha + \beta + k - j}{2j}G_{j - 2}^k
    - \dfrac{2\alpha + \beta + 2k - 3j}{2j}G_{j - 1}^k

@@

\textbf{TODO: Recursion for k > 1.\\}

@@

@@

    ?

>>>>

%
\section{The Spot Expansion}
\label{sec:spot}
%
\subsection{Functional form}
%
We adopt the following expression for the spherical harmonic coefficient
of degree $l$ and order $m$ in the expansion of a spot at
$\theta = \varphi = 0$:
%
\begin{align}
    \label{eq:ylm0}
    y_{lm}(r, \delta) =
    \begin{cases}
        1 - \dfrac{\delta r'}{2 (1 + r')}
         & l = m = 0    \\[2em]
        -\dfrac{\delta r' \left( 2 + r' \right)}
        {2 \sqrt{2l + 1} (1 + r')^{l + 1}}
         & l > 0, m = 0 \\[2em]
        0
         & m \ne 0
    \end{cases}
\end{align}
%
where
%
\begin{align}
    \label{eq:rprime}
    r' \equiv c_0 + c_1 r
\end{align}
%
and $c_0, c_1 > 0$ are scaling constants (defined in more detail below).
In the expressions above,
$\delta \in (-\infty, 1]$ is the spot contrast (the fractional
decrease in the brightness
at the center of the spot) and $r \in [0, 1]$ is the normalized spot radius.
The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
five important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity is azimuthally symmetric about the spot center
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $1 - \delta$
    \item The surface intensity at the antipode of the spot center is unity
    \item The size of the spot increases monotonically with $r$
\end{enumerate}
%
These properties may be demonstrated by considering the expression for the
surface intensity at a given point $(\theta, \varphi)$:
%
\begin{align}
    \begin{split}
        I(\theta, \varphi)
        & =
        \sum\limits_{l=0}^\infty \sum\limits_{m=-l}^l
        y_{lm} Y_{lm}(\theta, \varphi) \\
        & = \sum\limits_{l=0}^\infty
        y_{l0} \sqrt{2l + 1} P_l(\cos\theta)
    \end{split}
\end{align}
%
where $P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Note that the expression above is independent of the azimuth $\varphi$,
a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
expansion is therefore azimuthally symmetric about the spot center, as
stated in \textbf{1}.
%
Combining the above expression with Equation~(\ref{eq:ylm0}) and
rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    I(\theta) =
    1 + \dfrac{\delta r'}{2}
    -
    \dfrac{\delta r' \left( 2 + r' \right)}{2 (1 + r')}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + r'}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
    %
    \\
    \intertext{where}
    %
    \begin{split}
        A & = 1 + \dfrac{\delta r'}{2}                      \\
        B & = \delta r' (2 + r') \sqrt{\dfrac{1}{8 + 8 r'}} \\
        C & = \dfrac{1 + (1 + r')^2}{2 + 2 r'}
    \end{split}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{2}. The value at the minimum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
        & = 1 - \delta
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{3}, and the value at the maximum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
        & = 1
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{4}.
Finally, to show \textbf{5}, let us compute the half width at half minimum
$\Delta\theta$ of the intensity profile:
%
\begin{align}
    1 - \left(A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}}\right) =
    \dfrac{1}{2}\delta
\end{align}
%
Solving for $\Delta\theta$ yields
%
\begin{align}
    \label{eq:hwhm}
    \Delta\theta =
    \cos^{-1} \left[ \dfrac{2 + 3 r' (2 + r')}{2 (1 + r')^3} \right]
    \quad.
\end{align}
%
Differentiation with respect to $r'$ yields
%
\begin{align}
    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r'} =
    \frac{3 r' \left(2 + r'\right)}{\left(1 + r'\right)^4
        \sqrt{\frac{r'^2 \left(2 + r'\right)^2 \left(1 + 2 r'\right)
                \left(3 + 2 r'\right)}{\left(1 + r'\right)^6}}}
    \quad,
\end{align}
%
which is positive definite for all $r' > 0$.

\subsection{Behavior at finite degree}
%
The properties of the spot expansion described in the previous section
(in particular its monotonicity and values at the spot center and antipode)
are inherited from the generating function of the Legendre polynomials
(Equation~\ref{eq:gen}). These properties are therefore only rigorously
true when the expansion is taken to spherical
harmonic degree $l_{\mathrm{max}} = \infty$. For truncated expansions,
a value of $r'$ in Equation~(\ref{eq:ylm0}) that is too small will lead
to both ringing and a value at the spot center that is in general larger
than $1 - \delta$.

\begin{figure}[th!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/spot_expansion.pdf}
        \oscaption{spot_expansion}{%
            Polar intensity profiles for spots with different normalized
            radii $r$ in the range $(0, 1]$, computed at spherical harmonic
            degree $l_{\mathrm{max}} = 20$. Dotted black curves show profiles
            corresponding to $r < 0$ (corresponding to $r' < c_0$
            in Equation~\ref{eq:rprime}), for which the expansion does
            not converge, leading to ringing and a smaller spot contrast.
            \label{fig:spot_expansion}
        }
    \end{centering}
\end{figure}

To circumvent these issues, we impose a minimum radius
$r'_{\mathrm{min}}$, whose value is
a function of the degree of the expansion. Since the normalized spot radius
$r$ is defined in the range $[0, 1]$ (Equation~\ref{eq:rprime}), this
minimum value is equal to $c_0$. We compute $c_0 = c_0(l_{\mathrm{max}})$
by numerically finding the
value of $r'$ below which the fractional error in the intensity at the
center of the spot exceeds some threshold, which we take to be 0.01.
Since the width of the spot $\Delta\theta$ approaches $\pi$ as
$r' \rightarrow \infty$ (Equation~\ref{eq:hwhm}), we also impose a
maximum radius $r'_{\mathrm{max}}$, which we define to be the radius
corresponding to
$\Delta\theta = \nicefrac{5\pi}{12} = 75^\circ$. This choice is rather
arbitrary, but motivated by the fact that below this value, $\Delta\theta$
scales approximately logarithmically with $r'$ (see Figure~\ref{fig:hwhm}).

\begin{figure}[th!]
    \begin{centering}
        \includegraphics[width=0.4615\linewidth]{figures/hwhm.pdf}
        \includegraphics[width=0.475\linewidth]{figures/minr.pdf}
        \oscaption{minr}{%
            Foo bar.
            \codelink{hwhm}
            \label{fig:hwhm}
        }
    \end{centering}
\end{figure}

@@

        \dfrac{(\alpha + \beta + k - j)(1 + c_0)}{j(1 + c_0 + c_1)}
        \bigg] G_{j - 2}^k
    \nonumber
    + \hspace{1.5pt}
        -\dfrac{(\alpha + \beta - 2j + k)(1 + c_0) + (\alpha - j + k)c_1}
        {j(1 + c_0 + c_1)}
    G_{j - 1}^k
    \\[1.5em]
        \frac{(\alpha + \beta + k - 2)(1 + c_0)}
        {\lambda_{k - 1}(\alpha + \beta - j + k - 2)c_1}
    G_{j}^{k - 2}
    \nonumber
          & + \hspace{1.5pt}
        -\frac{\alpha - (\alpha + k)c_1 + \beta + k + (\alpha + \beta + k - 2)c_0 + (j + 2)c_1 - 2}
        {\lambda_{k-1} (\alpha + \beta - j + k - 2)c_1}
    G_{j}^{k - 1}

@@

@@

                \parbox[t]{\linewidth}{\vspace{-6em}$l > 0,        \\l' > 0, \\m = m' = 0$}

@@

                1 + \dfrac{\gamma_1}{1+c_0}
                \dfrac{\gamma_2}{2\gamma_1(1+c_0)}
                \dfrac{\gamma_1}{2\sqrt{2l + 1}(1+c_0)^{l+1}}
                - \dfrac{\gamma_2}{2\gamma_1(1 + c_0)} \bigg(
                \dfrac{\gamma_2}{4\sqrt{(2l + 1)(2l' + 1)}(1+c_0)^{l+l'+2}}
                \parbox[t]{\linewidth}{\vspace{-6em}$l > 0,        \\l' > 0, \\m = m' = 0$}
